### Manuscript: Myonuclear maturation dynamics in aged and adult regenerating mouse skeletal muscle

### This R notebook goes through the initial data processesing prior to conducting Monocle pseudotime analysis. In this script: 
#  1) Loading in cellraner ___ files and initial noise filtering using SoupX
#  2) Integrating data in Seurat
#  3) Dimensional reduction and cluster identification using Seurat

# Load required packages
```{r}
library(Seurat)
library(SoupX)
library(monocle)
library(cowplot)
library(ggplot2)
library(tidyr)
library(dplyr)
library(gplots)
library(limma)
library(gridExtra)
```

### Read in 10x files and filter with SoupX
```{r}
cellranger.dir <- "~/path_to_cellranger_outputs/cellranger_outputs/"

# Adult uninjured
AU <- load10X(paste(cellranger.dir, "AU/outs/", sep = ""))
count = autoEstCont(AU)
out = adjustCounts(count)
AU.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

# Old uninjured
OU <- load10X(paste(cellranger.dir, "OU/outs/", sep = ""))
count = autoEstCont(OU)
out = adjustCounts(count)
OU.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

# Adult 4 dpi
Aday4 <- load10X(paste(cellranger.dir, "Aday4/outs/", sep = ""))
count = autoEstCont(Aday4, tfidfMin = 0.5, forceAccept = T)
out = adjustCounts(count)
Aday4.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

# Adult 7 dpi
Aday7 <- load10X(paste(cellranger.dir, "Aday7/outs/", sep = ""))
count = autoEstCont(Aday7, tfidfMin = 0.5, forceAccept = T)
out = adjustCounts(count)
Aday7.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

# Old 4 dpi
Oday4 <- load10X(paste(cellranger.dir, "Oday4/outs/", sep = ""))
count = autoEstCont(Oday4,tfidfMin = 0.5, forceAccept = T)
out = adjustCounts(count)
Oday4.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

# Old 7 dpi
Oday7 <- load10X(paste(cellranger.dir, "Oday7/outs/", sep = ""))
count = autoEstCont(Oday7,tfidfMin = 0.5, forceAccept = T)
out = adjustCounts(count)
Oday7.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)
```

### Adding metadata before integrating seurat objects
```{r}
AU.obj@meta.data$cell.id <- 'AU'
OU.obj@meta.data$cell.id <- 'OU'
Aday4.obj@meta.data$cell.id <- 'Aday4'
Aday7.obj@meta.data$cell.id <- 'Aday7'
Oday4.obj@meta.data$cell.id <- 'Oday4'
Oday7.obj@meta.data$cell.id <- 'Oday7'

AU.obj@meta.data$condition.id <- 'adult'
OU.obj@meta.data$condition.id <- 'aged'
Aday4.obj@meta.data$condition.id <- 'adult'
Aday7.obj@meta.data$condition.id <- 'adult'
Oday4.obj@meta.data$condition.id <- 'aged'
Oday7.obj@meta.data$condition.id <- 'aged'
```

### Pre-cleaning violin QC plots
```{r}
AU.obj[["percent.mt"]] <- PercentageFeatureSet(AU.obj, pattern = "mt-")
VlnPlot(AU.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

OU.obj[["percent.mt"]] <- PercentageFeatureSet(OU.obj, pattern = "mt-")
VlnPlot(OU.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

Oday4.obj[["percent.mt"]] <- PercentageFeatureSet(Oday4.obj, pattern = "mt-")
VlnPlot(Oday4.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

Aday4.obj[["percent.mt"]] <- PercentageFeatureSet(Aday4.obj, pattern = "mt-")
VlnPlot(Aday4.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

Aday7.obj[["percent.mt"]] <- PercentageFeatureSet(Aday7.obj, pattern = "mt-")
VlnPlot(Aday7.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

Oday7.obj[["percent.mt"]] <- PercentageFeatureSet(Oday7.obj, pattern = "mt-")
VlnPlot(Oday7.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

### Cleaning data
```{r}
AU.obj <- subset(AU.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
OU.obj <- subset(OU.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
Aday4.obj <- subset(Aday4.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
Oday4.obj <- subset(Oday4.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
Aday7.obj <- subset(Aday7.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 3000 & percent.mt < 2) 
Oday7.obj <- subset(Oday7.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < .2) 
```

### Post-cleaning violin QC plots
```{r}
VlnPlot(AU.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(OU.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Oday4.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Aday4.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Aday7.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Oday7.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

### Merging before integrating
```{r}
all.nuclear.merged <- merge(AU.obj, OU.obj)
all.nuclear.merged <- merge(all.nuclear.merged, Aday4.obj)
all.nuclear.merged <- merge(all.nuclear.merged, Oday4.obj)
all.nuclear.merged <- merge(all.nuclear.merged, Aday7.obj)
all.nuclear.merged <- merge(all.nuclear.merged, Oday7.obj)
```

### Integration anchors
```{r}
list1 <- SplitObject(all.nuclear.merged, split.by = "condition.id")

anchors1 <- FindIntegrationAnchors(list1, dims = 1:20, anchor.features = 20000)

all.nuclear.integrated <- IntegrateData(anchors1, dims = 1:20 )
```

### All nuclei data combined (NOT subsetted for muscle nuclei)
```{r}
umap.function.integrated <- function(x, dims, n.neighbors, min.dist, resolution) 
    {
    DefaultAssay(x) <- "integrated"
    x <- ScaleData(x, verbose = FALSE)
    x <- RunPCA(x, npcs = 30, verbose = FALSE)
    x <- RunUMAP(x, reduction = "pca",  n.neighbors = 5, min.dist = .5, dims = dims) # Add these arguments into function up top? 
    x <- FindNeighbors(x, reduction = "pca", dims = dims)
    x <- FindClusters(x, resolution = 0.5)
    #DefaultAssay(x) <- "RNA"
}

all.nuclear.integrated.umap <- umap.function.integrated(all.nuclear.integrated , n.neighbors = 1, min.dist = .09, dims = 1:30, resolution = 8)

genelist <- c("Pax7", "Ttn", 'Tmem38a', 'Myh1', "Myh2", "Myh4", "Myh3", "Myh9", "Ache", "Col22a1", 'Pdgfra','Cacna1c', "Adgrf5")
DimPlot(all.nuclear.integrated.umap, pt.size = 2, label = T)
DimPlot(all.nuclear.integrated.umap, pt.size = 2, group.by = 'cell.id')
FeaturePlot(all.nuclear.integrated.umap, features = genelist,  pt.size = 2)
```

### Naming clusters 
```{r}
all.nuclear.integrated.umap.named <- all.nuclear.integrated.umap

new.cluster.ids <- c("Myogenic", #0
                     "Myogenic", #1
                     "Myogenic", #2
                     "Myogenic", #3
                     "Myogenic", #4
                     "Myogenic",#5
                     "Myogenic",#6
                     "FAPs/immune", #7
                     "Myogenic", #8
                     "Smooth Muscle", #9
                     "Endothelial" ,#10
                     "Myogenic", #11
                     "Myogenic", #12
                     "Myogenic") #13

names(new.cluster.ids) <- levels(all.nuclear.integrated.umap.named)
all.nuclear.integrated.umap.named <- RenameIdents(all.nuclear.integrated.umap.named, new.cluster.ids)

col.vec <- c("firebrick", "darkgoldenrod1", "forestgreen", "blue", "chocolate1", "black", "gray48")


DimPlot(all.nuclear.integrated.umap.named, pt.size = 2, label = F, cols = col.vec)
```

# UMAP plotting by condition.id
```{r}
DimPlot(all.nuclear.integrated.umap.named, pt.size = .5, label = F, group.by = 'condition.id', cols = c("black", "red"))
```

# UMAP plotting by cell.id
```{r}
DimPlot(all.nuclear.integrated.umap.named, pt.size = .5, label = F, group.by = 'cell.id', cols = c("black", "red"))
```

### Filtering out non-myogenic nuclei from original files & UMAP
```{r}
muscle.cells.1 <- subset(AU.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.2 <- subset(OU.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.3 <- subset(Aday4.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.4 <- subset(Oday4.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.5 <- subset(Aday7.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.6 <- subset(Oday7.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)

all.nuclear.subset.seurat <- merge(muscle.cells.1, muscle.cells.2)
all.nuclear.subset.seurat <- merge(all.nuclear.subset.seurat, muscle.cells.3)
all.nuclear.subset.seurat <- merge(all.nuclear.subset.seurat, muscle.cells.4)
all.nuclear.subset.seurat <- merge(all.nuclear.subset.seurat, muscle.cells.5)
all.nuclear.subset.seurat <- merge(all.nuclear.subset.seurat, muscle.cells.6)

list1 <- SplitObject(all.nuclear.subset.seurat, split.by = "condition.id")

anchors1 <- FindIntegrationAnchors(list1, dims = 1:20, anchor.features = 20000)

nuclear.subset.integrated <- IntegrateData(anchors1, dims = 1:20 )

umap.function.integrated <- function(x, dims, n.neighbors, min.dist, resolution) 
    {
    DefaultAssay(x) <- "integrated"
    x <- ScaleData(x, verbose = FALSE)
    x <- RunPCA(x, npcs = 30, verbose = FALSE)
    x <- RunUMAP(x, reduction = "pca",  n.neighbors = 5, min.dist = .5, dims = dims) # Add these arguments into function up top? 
    x <- FindNeighbors(x, reduction = "pca", dims = dims)
    x <- FindClusters(x, resolution = 0.5)
    #DefaultAssay(x) <- "RNA"
}

nuclear.subset.integrated.umap <- umap.function.integrated(nuclear.subset.integrated , n.neighbors = 1, min.dist = .09, dims = 1:30, resolution = 8)


DimPlot(nuclear.subset.integrated.umap, pt.size = .5, label = F, group.by = 'condition.id', cols = c("black", "red"))
```
