### Manuscript: Myonuclear maturation dynamics in aged and adult regenerating mouse skeletal muscle

### This R notebook goes through the the generation of a pseudotime trajectory from a Seurat object and more in-dept analysis of the branches and slope calculation. 

### Monocle main function for generating pseudotime trajectories
```{r}
main.function <- function(seurat.merged.obj, cth)
{
      data <- as(as.matrix(seurat.merged.obj@assays$RNA@data), 'sparseMatrix') 
      pd <- new('AnnotatedDataFrame', data = seurat.merged.obj@meta.data)
      fData <- data.frame(gene_short_name = row.names(data), row.names = row.names(data))
      fd <- new('AnnotatedDataFrame', data = fData)
      monocle.obj <- newCellDataSet(data,phenoData = pd,
                                    featureData = fd,
                                    lowerDetectionLimit = 0.5,
                                    expressionFamily = negbinomial.size())

      monocle.obj <- classifyCells(monocle.obj, cth, 0.1)
      print(table(pData(monocle.obj)$CellType))
      
      monocle.obj <- detectGenes(monocle.obj, min_expr = 0.1)
      expressed_genes.monocle.obj <- row.names(subset(fData(monocle.obj), num_cells_expressed >= 5))

      monocle.obj <- estimateSizeFactors(monocle.obj)
      monocle.obj <- estimateDispersions(monocle.obj)

      marker_diff.monocle.obj <- markerDiffTable(monocle.obj[expressed_genes.monocle.obj,], 
                                                 cth, 
                                                 residualModelFormulaStr = "~num_genes_expressed")

      diff_test_res.monocle.obj <- differentialGeneTest(monocle.obj[expressed_genes.monocle.obj,],
                                                        fullModelFormulaStr = "~CellType")
      
      ordering_genes.monocle.obj <- row.names(subset(diff_test_res.monocle.obj, qval < 0.01))
      
      monocle.obj <- setOrderingFilter(monocle.obj, ordering_genes.monocle.obj)
      
      monocle.obj <- reduceDimension(monocle.obj, max_components = 3, num_dim = 3,
        norm_method = 'log',
        reduction_method = 'tSNE',
        residualModelFormulaStr = "~CellType + num_genes_expressed",
        verbose = T, 
        check_duplicates = F)
      
      monocle.obj <- clusterCells(monocle.obj, num_clusters = 2)

      monocle.obj <- reduceDimension(monocle.obj, 
                                     max_components = 2,
                                     method = 'DDRTree',
                                     auto_param_selection = F)
      
      monocle.obj <- orderCells(monocle.obj, reverse = TRUE)
      
      monocle.new.name <<- monocle.obj 
}
```

### Constructing a cth - this is before running the actual monocle function (next chunk)
```{r}
x <- processed.seurat.obj
data <- as(as.matrix(x@assays$RNA@data), 'sparseMatrix') 
pd <- new('AnnotatedDataFrame', data = x@meta.data)
fData <- data.frame(gene_short_name = row.names(data), row.names = row.names(data))
fd <- new('AnnotatedDataFrame', data = fData)
x <- newCellDataSet(data,phenoData = pd,featureData = fd,lowerDetectionLimit = 0.5,expressionFamily = negbinomial.size())

### These designations are essentially useless - Monocle just requires that something be inputed as a cth
pax7_id <- row.names(subset(fData(x), gene_short_name == "Pax7"))
myf5_id <- row.names(subset(fData(x), gene_short_name == "Myf5"))
myog_id <- row.names(subset(fData(x), gene_short_name == "Myog"))
myod_id <- row.names(subset(fData(x), gene_short_name == "Myod1"))
myh1_id <- row.names(subset(fData(x), gene_short_name == "Myh1"))
myh2_id <- row.names(subset(fData(x), gene_short_name == "Myh2")) 
myh4_id <- row.names(subset(fData(x), gene_short_name == "Myh4")) 
tmem_id <- row.names(subset(fData(x), gene_short_name == "Tmem38a")) 

cth <- newCellTypeHierarchy()
cth <- addCellType(cth, "MuSC", classify_func = function(x) { x[pax7_id,] > 0  | x[myf5_id,] > 0 })
cth <- addCellType(cth, "early", classify_func = function(x) { x[myod_id,] > 0 | x[myog_id,] > 0 })
cth <- addCellType(cth, "Myonuclei", classify_func = function(x){ x[myh1_id,] > 0 | x[myh4_id,] > 0 | x[tmem_id,] > 0})
```

### Running monocle function (from top of page)
```{r}
main.function(processed.seurat.obj, cth)
monocle.obj <- monocle.new.name
```

```{r}
gene.list <- c("Tmem38a", "Musk", "Ncam1", "Pax7", "Ttn", "Ache")
plot_cell_trajectory(monocle.obj, color_by = "CellType")
plot_cell_trajectory(monocle.obj, color_by = "cell.id") 
plot_cell_trajectory(monocle.obj, color_by = "Pseudotime")
plot_cell_trajectory(monocle.obj, markers = gene.list, show_branch_points = F, use_color_gradient = TRUE)
```

### Designating branches by State designations - These were determined by observing plots from previous chunks
```{r}
branch1.ind <- which(as.numeric(as.vector(monocle.obj@phenoData@data[['State']])) == 1)
branch2.ind <- which(as.numeric(as.vector(monocle.obj@phenoData@data[['State']])) > 51)
branch3.ind <- which(as.numeric(as.vector(monocle.obj@phenoData@data[['State']])) > 1 & as.numeric(as.vector(all.nuclear.subset.monocle@phenoData@data[['State']])) < 52)


branch1.df <- cbind(branch1.ind, rep("branch1", length(branch1.ind)))
branch2.df <- cbind(branch2.ind, rep("branch2", length(branch2.ind)))
branch3.df <- cbind(branch3.ind, rep("branch3", length(branch3.ind)))

colnames(branch1.df) <- c("ind", "nuc.type")
colnames(branch2.df) <- c("ind", "nuc.type")
colnames(branch3.df) <- c("ind", "nuc.type")

branch1.df <- as.data.frame(branch1.df)
branch2.df <- as.data.frame(branch2.df)
branch3.df <- as.data.frame(branch3.df)

branch1.2.3.df <- rbind(branch1.df, branch2.df, branch3.df)

branch1.2.3.df.sorted <- branch1.2.3.df[order(as.numeric(as.vector(branch1.2.3.df$ind))),]

processed.seurat.obj$nuc.type <- branch1.2.3.df.sorted$nuc.type
monocle.obj$nuc.type <- branch1.2.3.df.sorted$nuc.type
```

### Counting nuclei by branch 
```{r}
nuc.type.nuclei <- processed.seurat.obj[['nuc.type']]
condition.nuclei <- processed.seurat.obj[['condition.id']]
df <- data.frame(nuc.type.nuclei, condition.nuclei)
table(df)
```


### ### ### ### ### ### ### ### ### 
### FAP/myogenic progenitor analysis
```{r}
tmp1 <- which(as.numeric(as.vector(monocle.obj@phenoData@data[['Pseudotime']])) <= 5)
tmp2 <- which(as.numeric(as.vector(monocle.obj@phenoData@data[['Pseudotime']])) > 5)

tmp1.names <- rep("progenitors", length(tmp1))
tmp2.names <- rep("everythingelse", length(tmp2))

tmp1.df <- as.data.frame(cbind(tmp1, tmp1.names))
tmp2.df <- as.data.frame(cbind(tmp2, tmp2.names))

colnames(tmp1.df) <- c("val", "name")
colnames(tmp2.df) <- c("val", "name")

tmp1.merged <- as.data.frame(rbind(tmp1.df, tmp2.df))

tmp2.merged <- tmp1.merged[order(as.numeric(as.vector(tmp1.merged$val)), decreasing = FALSE),]

processed.seurat.obj$FAP.analysis <- tmp2.merged$name
monocle.obj@phenoData@data[["FAP.analysis"]] <- processed.seurat.obj$FAP.analysis

plot_cell_trajectory(monocle.obj, color_by = "Pseudotime", pt.size = 7, )
plot_cell_trajectory(monocle.obj, color_by = "FAP.analysis", pt.size = 7)
```

### Using FindMarkers from Seurat to identify genes enriched in FAP/myogenic progenitor nuclei
```{r}
processed.seurat.obj <- SetIdent(processed.seurat.obj, value = "FAP.analysis")

markers <- FindMarkers(processed.seurat.obj, ident.1 = "progenitors")

markers %>% top_n(100, wt = avg_logFC) 

table.markers <- markers[which(markers$p_val_adj < 0.05 & markers$avg_logFC > 0),]
```

### ### ### ### ### ### ### ### ### 
### lncRNA analysis

# Constructing dataframe of pseudotime of individual lncRNA expression per individual nuclei
```{r}
pseudotime.experssion.df.function <- function(gene)
{
    gene.name <- paste(gene, "$", sep = "")
    df <- cbind(monocle.obj$Pseudotime, monocle.obj$condition.id, monocle.obj$State, monocle.obj@assayData$exprs[grep(gene.name, row.names(monocle.obj)),]) 
    
    df <- as.data.frame(df)
    colnames(df) <- c("pseudotime",  'condition.id', "state", "expression")
    
    row.names(df) <- NULL
    
    df.scatter <- cbind(as.character(df$condition.id), as.numeric(as.vector(df$pseudotime)), as.numeric(as.vector(df$expression)))
    df.scatter <- df.scatter[which(df.scatter[,3] > 0),]  
    df.scatter <- df.scatter[which(as.numeric(as.vector(df.scatter[,2])) < 71),]
    colnames(df.scatter) <- c('conditionid', "pseudotime", "expression")
    
    df.scatter <<- as.data.frame(df.scatter)
}

pseudotime.experssion.df.function("Runx1")
runx1.df.scatter <- df.scatter

pseudotime.experssion.df.function("Ncam1")
Ncam1.df.scatter <- df.scatter

pseudotime.experssion.df.function("Myh3")
Myh3.df.scatter <- df.scatter

pseudotime.experssion.df.function("Myh8")
Myh8.df.scatter <- df.scatter

pseudotime.experssion.df.function("Mirg")
Mirg.df.scatter <- df.scatter

pseudotime.experssion.df.function("Meg3")
Meg3.df.scatter <- df.scatter

pseudotime.experssion.df.function("Rian")
Rian.df.scatter <- df.scatter
```

### Plotting
```{r}
pdf(file = "path/filename.pdf" ,width = 9, height = 6)

ggplot() + 
  geom_smooth(data = runx1.df.scatter, aes( x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression))), size = 4, color = "gray3") + 
  geom_smooth(data = Ncam1.df.scatter, aes( x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression))), size = 4, color = "gray3") + 
  geom_smooth(data = Myh3.df.scatter, aes( x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression))), size = 4, color = "gray3") + 
  geom_smooth(data = Myh8.df.scatter, aes( x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression))), size = 4, color = "gray3") + 
  geom_smooth(data = Mirg.df.scatter, aes( x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression))), size = 4, color = "deepskyblue") +
  xlab("Pseudotime") + ylab("Expression") + theme_classic()
dev.off()
```

```{r}
pdf(file = "path/filename.pdf" ,width = 9, height = 6)

ggplot() + 
  geom_smooth(data = Meg3.df.scatter, aes( x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression))), size = 4, color = "deepskyblue2") + 
  geom_smooth(data = Rian.df.scatter, aes( x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression))), size = 4, color = "dodgerblue") + 
  xlab("Pseudotime") + ylab("Expression") + theme_classic()
dev.off()
```



### ### ### ### ### ### ### ### ### 
### Slope analysis

### Calculating slopes for glued lines between branch 1 and branch 2 OR branch 1 and branch 3
```{r}
slope.function.glued <- function(gene)
{
    gene.name <- paste(gene, "$", sep = "")
    df <- cbind(processed.seurat.obj$pseudotime, processed.seurat.obj$condition.id, processed.seurat.obj$state.num, as.character(processed.seurat.obj$nuc.type), monocle.obj@assayData$exprs[grep(gene.name, row.names(monocle.obj)),]) 
    
    row.names(df) <- NULL
    
    df.tmp <- df[which(df[,5] > 0),]  
    
    df.short <- as.data.frame(df.tmp)
    colnames(df.short) <- c("pseudotime",  'condition.id', "state", "nuc.type", "expression")
   
    branch1.ind <- grep("branch1", df.short$nuc.type)
    branch2.ind <- grep("branch2", df.short$nuc.type)
    branch3.ind <- grep("branch3", df.short$nuc.type)
    
    branch1.2.ind.combined <- c(branch1.ind, branch2.ind)
    branch1.3.ind.combined <- c(branch1.ind, branch3.ind)
    
    branch1.2.slope.df.short <<- df.short[branch1.2.ind.combined,]
    branch1.3.slope.df.short <<- df.short[branch1.3.ind.combined,]
    
    print("branch1:2 slope.short")  
    print(summary(lm(as.numeric(as.vector(branch1.2.slope.df.short$pseudotime)) ~ as.numeric(as.vector(branch1.2.slope.df.short$expression))))$coefficients[2,1])
    
    print("branch1:3 slope.short")
    print(summary(lm(as.numeric(as.vector(branch1.3.slope.df.short$pseudotime)) ~ as.numeric(as.vector(branch1.3.slope.df.short$expression))))$coefficients[2,1])



    df.long <- as.data.frame(df)
    colnames(df.long) <- c("pseudotime",  'condition.id', "state", "nuc.type", "expression")
   
    branch1.ind <- grep("branch1", df.long$nuc.type)
    branch2.ind <- grep("branch2", df.long$nuc.type)
    branch3.ind <- grep("branch3", df.long$nuc.type)
    
    branch1.2.ind.combined <- c(branch1.ind, branch2.ind)
    branch1.3.ind.combined <- c(branch1.ind, branch3.ind)
    
    branch1.2.slope.df.long <<- df.long[branch1.2.ind.combined,]
    branch1.3.slope.df.long <<- df.long[branch1.3.ind.combined,]
    
    print("branch1:2 slope.long")  
    print(summary(lm(as.numeric(as.vector(branch1.2.slope.df.long$pseudotime)) ~ as.numeric(as.vector(branch1.2.slope.df.long$expression))))$coefficients[2,1])
    
    print("branch1:3 slope.long")
    print(summary(lm(as.numeric(as.vector(branch1.3.slope.df.long$pseudotime)) ~ as.numeric(as.vector(branch1.3.slope.df.long$expression))))$coefficients[2,1])
    }



slope.function.glued("Esrrb") # Run individually for each gene and assemble into a table manually
esrrb.glued.df1.short <- branch1.2.slope.df.short
esrrb.glued.df2.short <- branch1.3.slope.df.short
```

### Plotting sloped lines
```{r}
slope.function.glued("Fgf1")
glued.df1 <- branch1.2.slope.df
glued.df2 <- branch1.3.slope.df


ggplotdf1 <- glued.df1
ggplotdf2 <- glued.df2

ggplot() +
  geom_smooth(data = ggplotdf1, aes(x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression))), color = "firebrick") + 
  geom_smooth(data = ggplotdf2, aes(x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression))), color = "green")
```

### To make PDF
```{r}
gene.list <- c("Pax7", "Ncam1", "Ttn", "Neb", "Ckm", "Tmem38a", "Neb", "Tnni2",  "Fgf1", "Ptpn3", "Actn2", "Tbc1d4", "Atp1b1", "Rcan2", "Ppara", "Esrrb", "Mylk2", "Mylk4", "Sorbs2", "Prkag3", "Ampd1", "Myoz1", "Fmnl2","Pfkfb3", "Actn3", "Eya4", "Pde4d", "Gpd2", "Pvalb" , "Ppara")
gene.list <- unique(gene.list)

for (i in 1:length(gene.list))
{
  pseudotime.experssion.function.full(as.character(gene.list[i]))
  alpha.val = 0.5


  print(ggplot(df.scatter.export.full, aes(x = as.numeric(as.vector(pseudotime)), y = as.numeric(as.vector(expression)))) + 
  geom_point(data=subset(df.scatter.export.full,nuc.type == 'branch3'),color = branch.color.scheme[2], alpha = alpha.val) + 
  geom_smooth(data=subset(df.scatter.export.full,nuc.type == 'branch3'),color = branch.color.scheme[2], alpha = alpha.val) +
  geom_point(data=subset(df.scatter.export.full,nuc.type == 'branch2'),color = branch.color.scheme[1], alpha = alpha.val) +
  geom_smooth(data=subset(df.scatter.export.full,nuc.type == 'branch2'),color = branch.color.scheme[1], alpha = alpha.val) +
  geom_point(data=subset(df.scatter.export.full,nuc.type == 'branch1'),color = branch.color.scheme[3], alpha = alpha.val)  +
  geom_smooth(data=subset(df.scatter.export.full,nuc.type == 'branch1'),color = branch.color.scheme[3], alpha = alpha.val) +
   xlab("Pseudotime") + ylab("Expression") + ggtitle(as.character(as.character(gene.list[i]))) + theme_classic())
}
```

### Aged vs adult scatterplots of gene expression and pseudotime (Figure 5)
```{r}
Idents(object = processed.seurat.obj) = "condition.id"

all.gene.names.vec <- row.names(processed.seurat.obj)
all.cells.names.vec <- colnames(processed.seurat.obj)


cells.pseudotime.state.condition.df <- as.data.frame(cbind(all.cells.names.vec,
                                             processed.seurat.obj$pseudotime,
                                             as.character(processed.seurat.obj$condition.id)))
colnames(cells.pseudotime.state.condition.df) <- c("cell.names", "cell.pseudotime.val", "condition")


pseudotime.expression.list <- list()
for (i in 1:length(all.gene.names.vec))
        {
          genename <- all.gene.names.vec[i]
          expr <- FetchData(object = processed.seurat.obj, vars = genename)
          colnames(expr) <- "tmp1.genename"
          tmp1 <-  which(x = expr$tmp1.genename > 0)
          df.per.gene <- cells.pseudotime.state.condition.df[tmp1,]
          
          adult.pseudotime.mean <- mean(as.numeric(as.vector(df.per.gene$cell.pseudotime.val[which(df.per.gene$condition == "adult")])))
          aged.pseudotime.mean <- mean(as.numeric(as.vector(df.per.gene$cell.pseudotime.val[which(df.per.gene$condition == "aged")])))
          
          adult.avg.expression <- AverageExpression(processed.seurat.obj, features = genename, verbose = F)$RNA$adult
          aged.avg.expression <- AverageExpression(processed.seurat.obj, features = genename, verbose = F)$RNA$aged
            
          pseudotime.expression.list[[i]] <- c( as.character(genename), 
                                                adult.pseudotime.mean , 
                                                aged.pseudotime.mean ,
                                                adult.avg.expression ,
                                                aged.avg.expression)
                                                                    
          names(pseudotime.expression.list[[i]]) <- c("gene", "adult.pseudotime.mean", "aged.pseudotime.mean", "adult.avg.expression", "aged.avg.expression")
          
          timing.vec <- seq(from = 1, to = 21315, by = 3000)
          if(i %in% timing.vec){print(i)}                                       
}


pseudotime.expression.df <- data.frame()
for (i in 1:length(pseudotime.expression.list))
{
  pseudotime.expression.df[i,1:5] <- c(as.character(pseudotime.expression.list[[i]][1]),
                                       as.numeric(as.vector(pseudotime.expression.list[[i]][2:5])))
  
  timing.vec <- seq(from = 1, to = 21315, by = 3000)
  if(i %in% timing.vec){print(i)}
}

colnames(pseudotime.expression.df) <- c("gene", "adult.pseudotime.mean", "aged.pseudotime.mean", "adult.avg.expression", "aged.avg.expression")
```

### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### 
### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ###  HEATMAPS ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### ### 

### ### ### ### ### ### ### ### ### 
### Generating heatmap of progenitor genes
```{r}
pdf(file = "path/filename.pdf" ,width = 7, height = 15)

progenitor.branch.genes.small <- c( "Ncam1", "Pax7", "Meg3", "Lama2", "Myh3", "Myh8","Hs3st5", "Agbl1", "Rian", "Mirg", "Ttn", "Tmem38a", "Rbms3", "Kirrel3", "Peg3", "Runx1", "Neb",   "Phkg1", "Ebf1", "Col5a3", "Col3a1", "Pdgfra")

plot_genes_branched_heatmap(monocle.obj[progenitor.branch.genes.small,],num_clusters =  2,branch_point = 7, show_rownames = T) ### 3/26/21 - use this one now

dev.off()
```

### ### ### ### ### ### ### ### ### 
### Generating heatmap for FAP/myogenic population genes
```{r}
FAP.branch1.genes <- c("Col6a1","Col6a3","Col4a1","Notch3","Emp1", "Cxcl12",  "Mrc2", "Fstl1", "Ebf1", "Col5a3", "Col3a1", "Pdgfra", "Dlc1", "Abca8a", "Tshz2", "Ror1", "Rhoj", "Ebf2", "Hmcn2", "Fbn1", "Lhfp", "Nav1", "Sox5", "Dcn", "Gsn", "Slit3", "Pde3a", "Egfr", "Gulp1", "Cald1", "Tgfbr3", "Bicc1", "Hs6st3", "Vcam1")
FAP.branch1.genes <- unique(FAP.branch1.genes)


pdf( file = "path/filename.pdf", width = 12, height = 15)

plot_genes_branched_heatmap(monocle.obj[FAP.branch1.genes,],num_clusters =  2,branch_point = 7, show_rownames = T) 

dev.off()

```

### ### ### ### ### ### ### ### ### 
### Generating heatmap of myonuclear genes
```{r}
myonuclei.branches <- c("Pax7", "Ncam1","Myh1", "Myh4", "Pfkfb3", "Actn2", "Actn3", "Sorbs2", "Pvalb", "Tbc1d4", "Prkag3", "Atp1b1", "Ank2", "Ampd1", "Fgf1", "Phka1", "Pde4d", "Rcan2", "Tpm1", "Mylk4", "Eya4", "Gpd2", "Acss2", "Tmem233", "Lama2", "Ptpn3", "Fmnl2", "Esrrb","Myoz1", "Myoz3", "Tnni2", "Ttn",  "Neb")


pdf(file = "path/filename.pdf" ,width = 10, height = 15)

plot_genes_branched_heatmap(monocle.obj[myonuclei.branches,],num_clusters =  3,branch_point = 28, show_rownames = T) 

dev.off()
```
