#### This notebook goes through:
# 1) Pacakge loading
# 2) loading in cellranger output files and preprocessing them for noise with soupx
# 3) Turn into seurat object and filter out low quality nuclei and doublets 


### 1) Pacakge loading
```{r}
library(Seurat)
library(SoupX)
library(monocle)
library(cowplot)
library(ggplot2)
library(tidyr)
library(dplyr)
library(gplots)
library(limma)
library(gridExtra)
```

### 2) loading in cellranger output files and preprocessing them for noise with soupx and creating seurat objects from them 
## "A" stands for "adult" and "O" stands for "Old" or "aged" as used for everything below
```{r}
input_directory <- "~/Desktop/Olwin_lab/snRNAseq/cellranger_outputs/"
output_directory 

AU <- load10X(paste(input_directory, "AU/outs/", sep=""))
count = autoEstCont(AU)
out = adjustCounts(count)
AU.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

OU <- load10X(paste(input_directory, "OU/outs/", sep=""))
count = autoEstCont(OU)
out = adjustCounts(count)
OU.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

Aday4 <- load10X(paste(input_directory, "Aday4/outs/", sep=""))
count = autoEstCont(Aday4, tfidfMin = 0.5, forceAccept = T)
out = adjustCounts(count)
Aday4.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

Aday7 <- load10X(paste(input_directory, "Aday7/outs/", sep=""))
count = autoEstCont(Aday7, tfidfMin = 0.5, forceAccept = T)
out = adjustCounts(count)
Aday7.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

Oday4 <- load10X(paste(input_directory, "Oday4/outs/", sep=""))
count = autoEstCont(Oday4,tfidfMin = 0.5, forceAccept = T)
out = adjustCounts(count)
Oday4.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)

Oday7 <- load10X(paste(input_directory, "Oday7/outs/", sep=""))
count = autoEstCont(Oday7,tfidfMin = 0.5, forceAccept = T)
out = adjustCounts(count)
Oday7.obj <- CreateSeuratObject(counts = out, project = "snRNAseq", min.cells = 3, min.features = 200)
```

### Adding metadata to seurat objects
```{r}
AU.obj@meta.data$cell.id <- 'AU'
OU.obj@meta.data$cell.id <- 'OU'
Aday4.obj@meta.data$cell.id <- 'Aday4'
Aday7.obj@meta.data$cell.id <- 'Aday7'
Oday4.obj@meta.data$cell.id <- 'Oday4'
Oday7.obj@meta.data$cell.id <- 'Oday7'

AU.obj@meta.data$condition.id <- 'Adult'
OU.obj@meta.data$condition.id <- 'Aged'
Aday4.obj@meta.data$condition.id <- 'Adult'
Aday7.obj@meta.data$condition.id <- 'Adult'
Oday4.obj@meta.data$condition.id <- 'Aged'
Oday7.obj@meta.data$condition.id <- 'Aged'
```

### Violin QC plots prior to filtering
```{r}
AU.obj[["percent.mt"]] <- PercentageFeatureSet(AU.obj, pattern = "mt-")
VlnPlot(AU.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

OU.obj[["percent.mt"]] <- PercentageFeatureSet(OU.obj, pattern = "mt-")
VlnPlot(OU.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

Oday4.obj[["percent.mt"]] <- PercentageFeatureSet(Oday4.obj, pattern = "mt-")
VlnPlot(Oday4.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

Aday4.obj[["percent.mt"]] <- PercentageFeatureSet(Aday4.obj, pattern = "mt-")
VlnPlot(Aday4.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

Aday7.obj[["percent.mt"]] <- PercentageFeatureSet(Aday7.obj, pattern = "mt-")
VlnPlot(Aday7.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

Oday7.obj[["percent.mt"]] <- PercentageFeatureSet(Oday7.obj, pattern = "mt-")
VlnPlot(Oday7.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

### Cleaning/filtering data
```{r}
AU.obj <- subset(AU.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
OU.obj <- subset(OU.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
Aday4.obj <- subset(Aday4.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
Oday4.obj <- subset(Oday4.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
Aday7.obj <- subset(Aday7.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 3000 & percent.mt < 2) # Changed nFeature_RNA and percent.mt
Oday7.obj <- subset(Oday7.obj, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < .2) # Changed percent.mt -- also use 0.2
```

### Post-filtering violin QC plots
```{r}
VlnPlot(AU.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(OU.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Oday4.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Aday4.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Aday7.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Oday7.obj, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

### Integrating all experiments together - first involves merging, then using the standard integration workflow in seurat

## Merging
```{r}
all.nuclear.merged <- merge(AU.obj, OU.obj)
all.nuclear.merged <- merge(all.nuclear.merged, Aday4.obj)
all.nuclear.merged <- merge(all.nuclear.merged, Oday4.obj)
all.nuclear.merged <- merge(all.nuclear.merged, Aday7.obj)
all.nuclear.merged <- merge(all.nuclear.merged, Oday7.obj)
```

## Integration anchors
```{r}
list1 <- SplitObject(all.nuclear.merged, split.by = "condition.id")

anchors1 <- FindIntegrationAnchors(list1, dims = 1:20, anchor.features = 20000)

all.nuclear.integrated <- IntegrateData(anchors1, dims = 1:20 )
```

## All nuclei data combined (NOT subsetted for muscle nuclei)
```{r}
umap.function.integrated <- function(x, dims, n.neighbors, min.dist, resolution) 
    {
    DefaultAssay(x) <- "integrated"
    x <- ScaleData(x, verbose = FALSE)
    x <- RunPCA(x, npcs = 30, verbose = FALSE)
    x <- RunUMAP(x, reduction = "pca",  n.neighbors = 5, min.dist = .5, dims = dims) 
    x <- FindNeighbors(x, reduction = "pca", dims = dims)
    x <- FindClusters(x, resolution = 0.5)
}

all.nuclear.integrated.umap <- umap.function.integrated(all.nuclear.integrated , n.neighbors = 1, min.dist = .09, dims = 1:30, resolution = 8)



genelist <- c("Pax7", "Ttn", 'Tmem38a', 'Myh1', "Myh2", "Myh4", "Myh3", "Myh9", "Ache", "Col22a1", 'Pdgfra','Cacna1c' , "Adgrf5")
DimPlot(all.nuclear.integrated.umap, pt.size = 2, label = T)
DimPlot(all.nuclear.integrated.umap, pt.size = 2, group.by = 'cell.id')
FeaturePlot(all.nuclear.integrated.umap, features = genelist,  pt.size = 2)
```


## Naming nuclear clusters by their identity as either myogenic or other - 4 clusters total
```{r}
all.nuclear.integrated.umap.named <- all.nuclear.integrated.umap

new.cluster.ids <- c("Myogenic", #0
                     "Myogenic", #1
                     "Myogenic", #2
                     "Myogenic", #3
                     "Myogenic", #4
                     "Myogenic",#5
                     "Myogenic",#6
                     "FAPs/immune", #7
                     "Myogenic", #8
                     "Smooth Muscle", #9
                     "Endothelial" ,#10
                     "Myogenic", #11
                     "Myogenic", #12
                     "Myogenic") #13

names(new.cluster.ids) <- levels(all.nuclear.integrated.umap.named)
all.nuclear.integrated.umap.named <- RenameIdents(all.nuclear.integrated.umap.named, new.cluster.ids)

col.vec <- c("firebrick", "darkgoldenrod1", "forestgreen", "blue", "chocolate1", "black", "gray48")

DimPlot(all.nuclear.integrated.umap.named, pt.size = 2, label = F, cols = col.vec)
```

## Plotting umap clusters by aged or adult
```{r}
DimPlot(all.nuclear.integrated.umap.named, pt.size = .5, label = F, group.by = 'condition.id', cols = c("black", "red"))
```


### ### ### ### Filtering out non-myogenic nuclei from original files & running monocle ### ### ### ### ### 

## Filtering out non-myogenic nuclei and UMAP
```{r}
muscle.cells.1 <- subset(AU.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.2 <- subset(OU.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.3 <- subset(Aday4.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.4 <- subset(Oday4.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.5 <- subset(Aday7.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)
muscle.cells.6 <- subset(Oday7.obj, Pax7 > 0 | Myod1 > 0 | Myog > 0 | Ckm > 0 |Mylk2 > 0)


all.nuclear.subset.seurat <- merge(muscle.cells.1, muscle.cells.2)
all.nuclear.subset.seurat <- merge(all.nuclear.subset.seurat, muscle.cells.3)
all.nuclear.subset.seurat <- merge(all.nuclear.subset.seurat, muscle.cells.4)
all.nuclear.subset.seurat <- merge(all.nuclear.subset.seurat, muscle.cells.5)
all.nuclear.subset.seurat <- merge(all.nuclear.subset.seurat, muscle.cells.6)

umap.function.standard <- function(x, dims) 
    {
    x <- NormalizeData(x, normalization.method = "LogNormalize", scale.factor = 10000)
    x.rownames <- rownames(x)
    DefaultAssay(x) <- "RNA"
    x <- ScaleData(x, features = x.rownames, verbose = FALSE)
    x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
    x <- RunPCA(x, npcs = 30, verbose = FALSE)
    x <- RunUMAP(x, reduction = "pca",  n.neighbors = 5, min.dist = .5, dims = dims) # Add these arguments into function up top? 
    x <- FindNeighbors(x, reduction = "pca", dims = dims)
    x <- FindClusters(x, resolution = 0.5)
}

all.nuclear.subset.umap <- umap.function.standard(all.nuclear.subset.seurat, dims = 1:20)

DimPlot(all.nuclear.subset.umap, pt.size = 2, label = T)
DimPlot(all.nuclear.subset.umap, pt.size = 2, group.by = 'cell.id')

```

## Making a celltype hierachy (cth) for Monocle - I don't really end up using this at all, its just a necessary object for monocle pipeline
```{r}
x <- all.nuclear.subset.umap
data <- as(as.matrix(x@assays$RNA@data), 'sparseMatrix') 
pd <- new('AnnotatedDataFrame', data = x@meta.data)
fData <- data.frame(gene_short_name = row.names(data), row.names = row.names(data))
fd <- new('AnnotatedDataFrame', data = fData)
x <- newCellDataSet(data,phenoData = pd,featureData = fd,lowerDetectionLimit = 0.5,expressionFamily = negbinomial.size())

pax7_id <- row.names(subset(fData(x), gene_short_name == "Pax7"))
myf5_id <- row.names(subset(fData(x), gene_short_name == "Myf5"))
myog_id <- row.names(subset(fData(x), gene_short_name == "Myog"))
myod_id <- row.names(subset(fData(x), gene_short_name == "Myod1"))

myh1_id <- row.names(subset(fData(x), gene_short_name == "Myh1"))
myh2_id <- row.names(subset(fData(x), gene_short_name == "Myh2")) 
myh4_id <- row.names(subset(fData(x), gene_short_name == "Myh4")) 

tmem_id <- row.names(subset(fData(x), gene_short_name == "Tmem38a")) 


cth <- newCellTypeHierarchy()
cth <- addCellType(cth, "MuSC", classify_func = function(x) { x[pax7_id,] > 0  | x[myf5_id,] > 0 })
cth <- addCellType(cth, "early", classify_func = function(x) { x[myod_id,] > 0 | x[myog_id,] > 0 })
cth <- addCellType(cth, "Myonuclei", classify_func = function(x){ x[myh1_id,] > 0 | x[myh4_id,] > 0 | x[tmem_id,] > 0})
```

### Running monocle pipeline on seurat object and cth to create trajectory 
```{r}
monocle.function <- function(seurat.merged.obj, cth)
{
      data <- as(as.matrix(seurat.merged.obj@assays$RNA@data), 'sparseMatrix') 
      pd <- new('AnnotatedDataFrame', data = seurat.merged.obj@meta.data)
      fData <- data.frame(gene_short_name = row.names(data), row.names = row.names(data))
      fd <- new('AnnotatedDataFrame', data = fData)
      monocle.obj <- newCellDataSet(data,phenoData = pd,
                                    featureData = fd,
                                    lowerDetectionLimit = 0.5,
                                    expressionFamily = negbinomial.size())

      monocle.obj <- classifyCells(monocle.obj, cth, 0.1)
      print(table(pData(monocle.obj)$CellType))
      
      monocle.obj <- detectGenes(monocle.obj, min_expr = 0.1)
      expressed_genes.monocle.obj <- row.names(subset(fData(monocle.obj), num_cells_expressed >= 5))

      monocle.obj <- estimateSizeFactors(monocle.obj)
      monocle.obj <- estimateDispersions(monocle.obj)

      marker_diff.monocle.obj <- markerDiffTable(monocle.obj[expressed_genes.monocle.obj,], 
                                                 cth, 
                                                 residualModelFormulaStr = "~num_genes_expressed")

      diff_test_res.monocle.obj <- differentialGeneTest(monocle.obj[expressed_genes.monocle.obj,],
                                                        fullModelFormulaStr = "~CellType")
      
      ordering_genes.monocle.obj <- row.names(subset(diff_test_res.monocle.obj, qval < 0.01))
      
      monocle.obj <- setOrderingFilter(monocle.obj, ordering_genes.monocle.obj)
      
      monocle.obj <- reduceDimension(monocle.obj, max_components = 3, num_dim = 3,
        norm_method = 'log',
        reduction_method = 'tSNE',
        residualModelFormulaStr = "~CellType + num_genes_expressed",
        verbose = T, 
        check_duplicates = F)
      
      monocle.obj <- clusterCells(monocle.obj, num_clusters = 2)

      monocle.obj <- reduceDimension(monocle.obj, 
                                     max_components = 2,
                                     method = 'DDRTree',
                                     auto_param_selection = F)
      
      monocle.obj <- orderCells(monocle.obj, reverse = TRUE)
      
      monocle.new.name <<- monocle.obj 
}

monocle.function(all.nuclear.subset.umap, cth)

all.nuclear.subset.monocle <- monocle.new.name
```

## Plotting trajectory with a few marker genes
```{r}
gene.list <- c("Tmem38a", "Musk", "Ncam1", "Pax7", "Ttn", "Ache")
plot_cell_trajectory(all.nuclear.subset.monocle, color_by = "CellType")
plot_cell_trajectory(all.nuclear.subset.monocle, color_by = "cell.id") 
plot_cell_trajectory(all.nuclear.subset.monocle, color_by = "Pseudotime")
plot_cell_trajectory(all.nuclear.subset.monocle, markers = gene.list, show_branch_points = F, use_color_gradient = TRUE)
```

































